---
title: "ABCD Flow of People"
author: "MochiBear.Hei"
date: "04.03.2025"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    page-layout: full
    html-math-method: katex
editor: source
---

```{css, echo = F}
body{
  font-family: Helvetica;
  font-size: 16pt;
  max-width: 1000px;
  margin: auto;
  margin-left:310px;
}
pre{
  font-size: 14px;
}
/* Headers */
h1{
    font-size: 24pt;
  }
h1,h2{
    font-size: 20pt;
  }
h3,h4,h5,h6{
  font-size: 18pt;
}
#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 300px;
  height: 100%;
  overflow:auto;
}
```

```{r setup,suppressPackageStartupMessages= TRUE}
#| include: false
#| message: false
#| warning: false
#| results: hide
knitr::opts_chunk$set(echo = TRUE, dpi = 300)

# Data manipulation
library(readr)
library(dplyr)
library(stringr)

# Plotting
library(ggplot2)
library(naniar)
library(tidyr)
library(gridExtra)

# Utilities
library(knitr)

```

Diagnosis save subject list
```{r}
subids <- read.csv("/Users/maggieheimvik/Documents/LCBC/Data/ABCD/core/mental-health/mh_p_ksads_ss.csv") %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  select(src_subject_id, ksads_11_917_p)   # Present diagnosis OCD F42

subids <- subids %>%
  filter(!is.na(ksads_11_917_p))
```
have mri
```{r}
mri <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/mri.csv") %>%
  select (-smri_vol_scs_lesionlh, -smri_vol_scs_lesionrh, -smri_vol_scs_intracranialv) #removed these columns because they have 4897 NA each

# merge
full_dat <- inner_join(mri, subids, by = "src_subject_id")
```

## Quality control of images 
```{r}
qc <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/core/imaging/mri_y_qc_incl.csv") %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  select(src_subject_id, imgincl_t1w_include) %>%
  distinct(src_subject_id, .keep_all = TRUE)

full_dat <- inner_join(full_dat, qc, by = "src_subject_id")
```

1 = include: The data is recommended for inclusion.
0 = exclude: The data does not meet the recommended criteria for inclusion.
```{r}
#remove those with imgincl_t1w_include = 0
full_dat <- full_dat %>%
  filter(imgincl_t1w_include == 1)

```

have bpm
```{r}
bpm <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/core/mental-health/mh_y_bpm.csv") %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  select(src_subject_id, bpm_y_scr_internal_t) %>%
  distinct(src_subject_id, .keep_all = TRUE) %>% #remove duplicates
  mutate(across(c(bpm_y_scr_internal_t), ~ as.numeric(scale(.)))) #standardize

# merge
full_dat <- inner_join(full_dat, bpm, by = "src_subject_id")
```

have cbcl
```{r}
cbcl <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/mental-health/mh_p_cbcl.csv") %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  select(
    src_subject_id,cbcl_scr_syn_internal_t, cbcl_scr_syn_totprob_t
  ) %>%
  distinct(src_subject_id, .keep_all = TRUE) %>% # remove duplicates
  mutate(across(c(cbcl_scr_syn_internal_t,  cbcl_scr_syn_totprob_t), ~ as.numeric(scale(.)))) # standardize scores

# merge
full_dat <- inner_join(full_dat, cbcl, by = "src_subject_id")
```

missingness
```{r}
na_counts <- colSums(is.na(full_dat))
na_counts[na_counts > 0]  # print only columns with NA values


# mean imputation: 
full_dat <- full_dat %>%
  mutate(across(c(mrisdp_508, mrisdp_527,    mrisdp_567, mrisdp_582, mrisdp_601, mrisdp_602, mrisdp_603, mrisdp_604),
                #  bpm_y_scr_internal_t,
                ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))

full_dat <- full_dat %>%
  filter(rowSums(is.na(select(., bpm_y_scr_internal_t))) == 0)
```

#medication
```{r}
antipsychotics <- c("Risperidone", "Aripiprazole", "Olanzapine", "Quetiapine", "Haloperidol", 
                    "Ziprasidone", "Clozapine", "Lurasidone", "Paliperidone")

antidepressants <- c("Fluoxetine", "Sertraline", "Citalopram", "Escitalopram", "Paroxetine", 
                     "Bupropion", "Duloxetine", "Venlafaxine", "Desvenlafaxine", "Amitriptyline")

# list of med columns in csv 
med_cols <- paste0("med", 1:15, "_rxnorm_p")

medication <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/core/physical-health/ph_p_meds.csv") %>%
  select(src_subject_id, eventname, all_of(med_cols)) %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  distinct(src_subject_id, .keep_all = TRUE)
```

```{r}
# make two new boolean columns 
medication$uses_antipsychotic <- apply(medication[, med_cols], 1, function(row) {
  any(sapply(antipsychotics, function(drug) grepl(drug, row, ignore.case = TRUE)))
})
medication$uses_antidepressant <- apply(medication[, med_cols], 1, function(row) {
  any(sapply(antidepressants, function(drug) grepl(drug, row, ignore.case = TRUE)))
})

#check
summary(medication$uses_antipsychotic)
summary(medication$uses_antidepressant)

#keep only the boolean columns 
medication <- medication %>%
  select(src_subject_id, uses_antipsychotic, uses_antidepressant)
```

remove those in analysis dataset using medication
```{r}
full_dat <- inner_join(full_dat, medication, by = "src_subject_id")

# Remove rows from full_dat where participants are using antipsychotics or antidepressants
full_dat <- full_dat[!(full_dat$uses_antipsychotic == TRUE | full_dat$uses_antidepressant == TRUE), ]
```

#refresh 
# OCD 0 or 1
```{r}
final_data <- full_dat %>%
  select(
    src_subject_id, ksads_11_917_p
  )
  
```


In healthy group keep only those that do not have any comorbidities
Healthy vs OCD
```{r}
# Load additional comorbidities data
all_ksads_diagnosis <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/comorbidities.csv") %>%
  select(
    src_subject_id,
    adhd,
    any_bipolar,
    any_conduct_odd,
    any_depressive_disorder,
    any_anxiety_disorder,
    any_drug_use_disorder,
    any_eating_disorder,
    ptsd,
    any_suicidality,
    no_diagnosis,
    only_ocd,
    comorbidities
  ) 

# merge
final_data <- inner_join(final_data, all_ksads_diagnosis, by = "src_subject_id")

# Filter out those with ksads_11_917_p = 0 and comorbidities > 0
final_data <- final_data %>%
  filter(!(ksads_11_917_p == 0 & comorbidities > 0))
# Count for people with OCD (ksads_11_917_p = 1)
ocd_summary <- final_data %>%
  filter(ksads_11_917_p == 1) %>%
  summarise(
    count_with_comorbidities = sum(comorbidities > 0),
    count_without_comorbidities = sum(comorbidities == 0)
  )

# Count for people without OCD (ksads_11_917_p = 0)
non_ocd_summary <- final_data %>%
  filter(ksads_11_917_p == 0) %>%
  summarise(
    count_with_comorbidities = sum(comorbidities > 0),
    count_without_comorbidities = sum(comorbidities == 0)
  )

# Print the summaries
cat("OCD Diagnosis Summary:\n")
cat("Count of OCD (ksads_11_917_p = 1) with comorbidities > 0:", ocd_summary$count_with_comorbidities, "\n")
cat("Count of OCD (ksads_11_917_p = 1) with comorbidities = 0:", ocd_summary$count_without_comorbidities, "\n")

cat("\nNon-OCD Diagnosis Summary:\n")
cat("Count of Non-OCD (ksads_11_917_p = 0) with comorbidities > 0:", non_ocd_summary$count_with_comorbidities, "\n")
cat("Count of Non-OCD (ksads_11_917_p = 0) with comorbidities = 0:", non_ocd_summary$count_without_comorbidities, "\n")
```


#save for analysis 
```{r}
final_data <- final_data %>%
  select(src_subject_id, ksads_11_917_p, no_diagnosis, only_ocd, comorbidities)


write.csv(final_data, "/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/subids.csv", row.names = FALSE)
```

#save for demogrpahics
```{r}
full_dat <- full_dat %>%
  select(
    src_subject_id
  )
write.csv(full_dat, "/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/demo_subid.csv", row.names = FALSE)
```


}
which src_subj are using meds
```{r}
#find
antipsychotic_users <- which(medication$uses_antipsychotic == TRUE)
antidepressant_users <- which(medication$uses_antidepressant == TRUE)

# View 
antipsychotic_details <- medication[antipsychotic_users, c("src_subject_id", med_cols)]
antidepressant_details <- medication[antidepressant_users, c("src_subject_id", med_cols)]

# Print 
print("Participants using antipsychotics:")
print(antipsychotic_details)

print("Participants using antidepressants:")
print(antidepressant_details)
```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

```{r}


```

