---
title: "Demographic characteristics"
author: "MochiBear.Hei"
date: "12.20.2024"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    page-layout: full
    html-math-method: katex
editor: source
---

```{css, echo = F}
body{
  font-family: Helvetica;
  font-size: 16pt;
  max-width: 1000px;
  margin: auto;
  margin-left:310px;
}
pre{
  font-size: 14px;
}
/* Headers */
h1{
    font-size: 24pt;
  }
h1,h2{
    font-size: 20pt;
  }
h3,h4,h5,h6{
  font-size: 18pt;
}
#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 300px;
  height: 100%;
  overflow:auto;
}
```

```{r setup,suppressPackageStartupMessages= TRUE}
#| include: false
#| message: false
#| warning: false
#| results: hide
knitr::opts_chunk$set(echo = TRUE, dpi = 300)

# Data manipulation
library(dplyr)
library(tidyr)

# Plotting
library(ggplot2)

# Utilities
library(knitr)

library(gt)
```


# Cleaning
## Diagnosis and ksads demo
```{r}
diagnosis <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/OCD/diagnosis.csv")

# Select only the specified columns
diagnosis <- diagnosis %>%
  select(src_subject_id, 
         ksads_11_917_p) #present diagnosis OCD F42

demo <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/ocd/demo.csv")

# Merge
## Demographic variables: sex, gender, ethnicity, age + diagnosis
demo <- diagnosis %>%
  left_join(demo, by = "src_subject_id")

```

```{r}
demo_2 <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/abcd-general/abcd_p_demo.csv") %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  mutate(
    demo_comb_income_v2_l = if_else(demo_comb_income_v2_l %in% c(777, 999), NA_real_, demo_comb_income_v2_l),
    demo_prnt_marital_v2_l = if_else(demo_prnt_marital_v2_l %in% c(777, 999), NA_real_, demo_prnt_marital_v2_l),
    demo_prtnr_ed_v2_2yr_l = if_else(demo_prtnr_ed_v2_2yr_l %in% c(777, 999), NA_real_, demo_prtnr_ed_v2_2yr_l)
  )

# Select only the specified columns
demo_2 <- demo_2 %>%
  select(src_subject_id, 
         demo_prnt_marital_v2_l, # Marital status of other parent
         demo_prnt_ed_v2_2yr_l, # Parental education level
         demo_prtnr_ed_v2_2yr_l, # Other parent education level 
         demo_comb_income_v2_l # Other parent income level
         ) 
#Merge
demo <- demo_2 %>%
  left_join(demo, by = "src_subject_id")%>%
  filter(!is.na(ksads_11_917_p))
```

```{r}
#how many in the full sample?
# Calculate the total number of unique src_subject_id
total_unique_subjects <- demo %>%
  summarise(total_unique_subjects = n_distinct(src_subject_id)) %>%
  pull(total_unique_subjects)

# Display the result
cat("Total number of unique src_subject_id:", total_unique_subjects, "\n")


## how many with no ocd
# Calculate the total number of people with OCD absent
total_ocd_absent <- demo %>%
  filter(ksads_11_917_p != 1) %>% # Filter for cases where OCD is absent
  summarise(total_ocd_absent = n_distinct(src_subject_id)) %>%
  pull(total_ocd_absent)
# Display the result
cat("Total number of people with OCD absent:", total_ocd_absent, "\n")

## how many with ocd
# Calculate the total number of people with OCD absent
total_ocd_present <- demo %>%
  filter(ksads_11_917_p == 1) %>% # Filter for cases where OCD is absent
  summarise(total_ocd_present = n_distinct(src_subject_id)) %>%
  pull(total_ocd_present)
# Display the result
cat("Total number of people with OCD present:", total_ocd_present, "\n")
```

Formatting variables
```{r}
str(demo)

# Convert variables to factor with meaningful labels
demo <- demo %>%
  mutate(
    demo_prnt_marital_v2_l = factor(demo_prnt_marital_v2_l, levels = c(1, 2, 3, 4, 5, 6, 777),
                                    labels = c("Married", "Widowed", "Divorced", "Separated", "Never married", "Living with partner", "Refused to answer")),
    
    demo_prnt_ed_v2_2yr_l = factor(demo_prnt_ed_v2_2yr_l, levels = c(0:14, 16:23, 777, 999),
                                    labels = c("Never attended/Kindergarten only", "1st grade", "2nd grade", 
                                               "3rd grade", "4th grade", "5th grade", "6th grade", 
                                               "7th grade", "8th grade", "9th grade", "10th grade", 
                                               "11th grade", "12th grade, no diploma", 
                                               "High school graduate", "GED or equivalent", 
                                               "Less than 1 year of college credit/post-secondary education", 
                                               "One year or more of college credit, no degree",
                                               "Associate degree: Occupational, Technical, or Vocational", 
                                               "Associate degree: Academic Program", 
                                               "Bachelor's degree", "Master's degree",
                                               "Professional School degree", "Doctoral degree", 
                                               "Refused to answer", "Don't Know")),
    
    demo_prtnr_ed_v2_2yr_l = factor(demo_prtnr_ed_v2_2yr_l, levels = c(0:14, 16:23, 777, 999),
                                    labels = c("Never attended/Kindergarten only", "1st grade", "2nd grade", 
                                               "3rd grade", "4th grade", "5th grade", "6th grade", 
                                               "7th grade", "8th grade", "9th grade", "10th grade", 
                                               "11th grade", "12th grade, no diploma", 
                                               "High school graduate", "GED or equivalent", 
                                               "Associate degree: Occupational, Technical, or Vocational",
                                               "Associate degree: Academic Program", 
                                               "Bachelor's degree", "Master's degree",
                                               "Professional School degree", "Doctoral degree", 
                                               "Less than 1 year of college credit/post-secondary education", 
                                               "One year or more of college credit, no degree",
                                               "Refused to answer", "Don't Know")),
    
    demo_comb_income_v2 = factor(demo_comb_income_v2_l),
    ksads_11_917_p = as.factor(ksads_11_917_p),  
    demo_sex_v2 = as.factor(demo_sex_v2),      
    demo_gender_id_v2 = as.factor(demo_gender_id_v2),  
    race_ethnicity = as.factor(race_ethnicity),  
    demo_brthdat_v2 = as.integer(demo_brthdat_v2) 
  )

# Check
str(demo)

```

## Counts and percents
```{r}
# Define the given totals
total_full_sample <- 10711
total_ocd_absent <- 10049
total_ocd_present <- 662

# Identify the OCD presence
demo <- demo %>%
  mutate(has_ocd = if_else(ksads_11_917_p == 1, "OCD", "No OCD"))

# Function to categorize and summarize demographic data
summarize_demographics <- function(data, demographic_var) {
  data %>%
    group_by(.data[[demographic_var]], has_ocd) %>%
    summarise(count = n(), .groups = "drop") %>%
    mutate(
      total = case_when(
        has_ocd == "OCD" ~ total_ocd_present,
        has_ocd == "No OCD" ~ total_ocd_absent,
        TRUE ~ total_full_sample
      ),
      percentage = (count / total) * 100
    )
}

# List of demographic variables to summarize
demographic_vars <- c("demo_prnt_marital_v2_l", "demo_prnt_ed_v2_2yr_l", 
                      "demo_prtnr_ed_v2_2yr_l", "demo_comb_income_v2",
                      "demo_sex_v2", "demo_gender_id_v2", "race_ethnicity",
                      "demo_brthdat_v2")

# Create a summary for each demographic variable
demographic_summaries <- lapply(demographic_vars, function(var) {
  summarize_demographics(demo, var)
})

# Print the summarized data
names(demographic_summaries) <- demographic_vars
lapply(seq_along(demographic_vars), function(i) {
  cat("\nSummary for", demographic_vars[i], ":\n")
  print(demographic_summaries[[i]])
})
```

## Mean and SD
```{r}

# Full sample statistics for numeric demographic variables
full_sample_stats <- demo %>%
  summarise(
    mean_age = mean(demo_brthdat_v2, na.rm = TRUE),
    sd_age = sd(demo_brthdat_v2, na.rm = TRUE),
    mean_income = mean(demo_comb_income_v2_l, na.rm = TRUE),
    sd_income = sd(demo_comb_income_v2_l, na.rm = TRUE),
    mean_prnt_ed = mean(as.numeric(demo_prnt_ed_v2_2yr_l), na.rm = TRUE),
    sd_prnt_ed = sd(as.numeric(demo_prnt_ed_v2_2yr_l), na.rm = TRUE),
    mean_prtnr_ed = mean(as.numeric(demo_prtnr_ed_v2_2yr_l), na.rm = TRUE),
    sd_prtnr_ed = sd(as.numeric(demo_prtnr_ed_v2_2yr_l), na.rm = TRUE)
  )

# OCD present statistics for numeric demographic variables
ocd_present_stats <- demo %>%
  filter(ksads_11_917_p == 1) %>%
  summarise(
    mean_age = mean(demo_brthdat_v2, na.rm = TRUE),
    sd_age = sd(demo_brthdat_v2, na.rm = TRUE),
    mean_income = mean(demo_comb_income_v2_l, na.rm = TRUE),
    sd_income = sd(demo_comb_income_v2_l, na.rm = TRUE),
    mean_prnt_ed = mean(as.numeric(demo_prnt_ed_v2_2yr_l), na.rm = TRUE),
    sd_prnt_ed = sd(as.numeric(demo_prnt_ed_v2_2yr_l), na.rm = TRUE),
    mean_prtnr_ed = mean(as.numeric(demo_prtnr_ed_v2_2yr_l), na.rm = TRUE),
    sd_prtnr_ed = sd(as.numeric(demo_prtnr_ed_v2_2yr_l), na.rm = TRUE)
  )

# OCD absent statistics for numeric demographic variables
ocd_absent_stats <- demo %>%
  filter(ksads_11_917_p == 0) %>%
  summarise(
    mean_age = mean(demo_brthdat_v2, na.rm = TRUE),
    sd_age = sd(demo_brthdat_v2, na.rm = TRUE),
    mean_income = mean(demo_comb_income_v2_l, na.rm = TRUE),
    sd_income = sd(demo_comb_income_v2_l, na.rm = TRUE),
    mean_prnt_ed = mean(as.numeric(demo_prnt_ed_v2_2yr_l), na.rm = TRUE),
    sd_prnt_ed = sd(as.numeric(demo_prnt_ed_v2_2yr_l), na.rm = TRUE),
    mean_prtnr_ed = mean(as.numeric(demo_prtnr_ed_v2_2yr_l), na.rm = TRUE),
    sd_prtnr_ed = sd(as.numeric(demo_prtnr_ed_v2_2yr_l), na.rm = TRUE)
  )

# Print summaries
cat("Full Sample Statistics\n")
print(full_sample_stats)

cat("\nOCD Present Statistics\n")
print(ocd_present_stats)

cat("\nOCD Absent Statistics\n")
print(ocd_absent_stats)

```


## Cognition
```{r}
cogn <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/neurocognition/nc_y_nihtb.csv")

# Select only the specified columns
cogn <- cogn %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  select(src_subject_id, 
         nihtbx_totalcomp_fc #NIH toolbox - cognition total
         ) 
```

## Pubertal status
Petersen, A. C., Crockett, L., et al. (1988) A self-report measure of pubertal status: Reliability, validity, and initial norms. J Youth Adolesc 17(2): 117-133.

puber <- puber %>%
  mutate(
    pds_p_ss_female_category = factor(pds_p_ss_female_category, levels = c(1, 2, 3, 4, 5),
                                    labels = c("Prepuberty", "Early Puberty", "Mid Puberty", "Late Puberty", "Post puberty"))
  )

```{r}
puber <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/physical-health/ph_p_pds.csv")

# Select only the specified columns
puber <- puber %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  select(src_subject_id, 
         pds_p_ss_female_category, # Pubertal status female
         pds_p_ss_female_category_2, #based on sex at birth
         pds_p_ss_male_category, # Pubertal status male
         pds_p_ss_male_category_2 #based on sex at birth
         ) 

diagnosis <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/OCD/diagnosis.csv")

# Select only the specified columns
diagnosis <- diagnosis %>%
  select(src_subject_id, 
         ksads_11_917_p) #present diagnosis OCD F42

# Merge the puberty and diagnosis data on src_subject_id
puber <- puber %>%
  left_join(diagnosis, by = "src_subject_id") 

#how many in the full sample?
# Calculate the total number of unique src_subject_id
total_unique_subjects <- puber %>%
  summarise(total_unique_subjects = n_distinct(src_subject_id)) %>%
  pull(total_unique_subjects)

# Display the result
cat("Total number of unique src_subject_id:", total_unique_subjects, "\n")


## how many with no ocd
# Calculate the total number of people with OCD absent
total_ocd_absent <- puber %>%
  filter(ksads_11_917_p != 1) %>% # Filter for cases where OCD is absent
  summarise(total_ocd_absent = n_distinct(src_subject_id)) %>%
  pull(total_ocd_absent)
# Display the result
cat("Total number of people with OCD absent:", total_ocd_absent, "\n")

## how many with ocd
# Calculate the total number of people with OCD absent
total_ocd_present <- puber %>%
  filter(ksads_11_917_p == 1) %>% # Filter for cases where OCD is absent
  summarise(total_ocd_present = n_distinct(src_subject_id)) %>%
  pull(total_ocd_present)
# Display the result
cat("Total number of people with OCD present:", total_ocd_present, "\n")
```


```{r}
# Define the given totals
total_full_sample <- 10973
total_ocd_absent <- 10068
total_ocd_present <- 662

# Full sample statistics for puberty variables
full_sample_stats_pds <- puber %>%
  summarise(
    mean_female_cat = mean(pds_p_ss_female_category, na.rm = TRUE),
    sd_female_cat = sd(pds_p_ss_female_category, na.rm = TRUE),
    mean_female_cat_2 = mean(pds_p_ss_female_category_2, na.rm = TRUE),
    sd_female_cat_2 = sd(pds_p_ss_female_category_2, na.rm = TRUE),
    mean_male_cat = mean(pds_p_ss_male_category, na.rm = TRUE),
    sd_male_cat = sd(pds_p_ss_male_category, na.rm = TRUE),
    mean_male_cat_2 = mean(pds_p_ss_male_category_2, na.rm = TRUE),
    sd_male_cat_2 = sd(pds_p_ss_male_category_2, na.rm = TRUE)
  )

# OCD present statistics for puberty variables
ocd_present_stats_pds <- puber %>%
  filter(ksads_11_917_p == 1) %>%
  summarise(
    mean_female_cat = mean(pds_p_ss_female_category, na.rm = TRUE),
    sd_female_cat = sd(pds_p_ss_female_category, na.rm = TRUE),
    mean_female_cat_2 = mean(pds_p_ss_female_category_2, na.rm = TRUE),
    sd_female_cat_2 = sd(pds_p_ss_female_category_2, na.rm = TRUE),
    mean_male_cat = mean(pds_p_ss_male_category, na.rm = TRUE),
    sd_male_cat = sd(pds_p_ss_male_category, na.rm = TRUE),
    mean_male_cat_2 = mean(pds_p_ss_male_category_2, na.rm = TRUE),
    sd_male_cat_2 = sd(pds_p_ss_male_category_2, na.rm = TRUE)
  )

# OCD absent statistics for puberty variables
ocd_absent_stats_pds <- puber %>%
  filter(ksads_11_917_p == 0) %>%
  summarise(
    mean_female_cat = mean(pds_p_ss_female_category, na.rm = TRUE),
    sd_female_cat = sd(pds_p_ss_female_category, na.rm = TRUE),
    mean_female_cat_2 = mean(pds_p_ss_female_category_2, na.rm = TRUE),
    sd_female_cat_2 = sd(pds_p_ss_female_category_2, na.rm = TRUE),
    mean_male_cat = mean(pds_p_ss_male_category, na.rm = TRUE),
    sd_male_cat = sd(pds_p_ss_male_category, na.rm = TRUE),
    mean_male_cat_2 = mean(pds_p_ss_male_category_2, na.rm = TRUE),
    sd_male_cat_2 = sd(pds_p_ss_male_category_2, na.rm = TRUE)
  )

# Print summaries
cat("Full Sample Stats for Puberty Variables\n")
print(full_sample_stats_pds)

cat("\nOCD Present Stats for Puberty Variables\n")
print(ocd_present_stats_pds)

cat("\nOCD Absent Stats for Puberty Variables\n")
print(ocd_absent_stats_pds)

```

## Grades
```{r}
grades <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/culture-environment/ce_p_sag.csv")

#Keep the relevant time points
grades <- grades %>%
  filter(eventname == "2_year_follow_up_y_arm_1")

# Select only the specified columns
grades <- grades %>%
  select(sag_grade_type)
```

## SES variables
o	SES can be reported using variables like parental education level, household income, or an SES index if available.

```{r}
latent <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/abcd-general/abcd_y_lf.csv")


diagnosis <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/OCD/diagnosis.csv")

# Select only the specified columns
diagnosis <- diagnosis %>%
  select(src_subject_id, 
         ksads_11_917_p) #present diagnosis OCD F42

# Merge the SES variables and diagnosis data on src_subject_id
latent <- latent %>%
  left_join(diagnosis, by = "src_subject_id")  %>%
  select(-eventname)


# Full sample statistics for latent variables
full_sample_stats_latent <- latent %>%
  summarise(
    mean_general_ses = mean(latent_factor_ss_general_ses, na.rm = TRUE),
    sd_general_ses = sd(latent_factor_ss_general_ses, na.rm = TRUE),
    mean_social = mean(latent_factor_ss_social, na.rm = TRUE),
    sd_social = sd(latent_factor_ss_social, na.rm = TRUE),
    mean_perinatal = mean(latent_factor_ss_perinatal, na.rm = TRUE),
    sd_perinatal = sd(latent_factor_ss_perinatal, na.rm = TRUE)
  )

# OCD present statistics for latent variables
ocd_present_stats_latent <- latent %>%
  filter(ksads_11_917_p == 1) %>%
  summarise(
    mean_general_ses = mean(latent_factor_ss_general_ses, na.rm = TRUE),
    sd_general_ses = sd(latent_factor_ss_general_ses, na.rm = TRUE),
    mean_social = mean(latent_factor_ss_social, na.rm = TRUE),
    sd_social = sd(latent_factor_ss_social, na.rm = TRUE),
    mean_perinatal = mean(latent_factor_ss_perinatal, na.rm = TRUE),
    sd_perinatal = sd(latent_factor_ss_perinatal, na.rm = TRUE)
  )

# OCD absent statistics for latent variables
ocd_absent_stats_latent <- latent %>%
  filter(ksads_11_917_p == 0) %>%
  summarise(
    mean_general_ses = mean(latent_factor_ss_general_ses, na.rm = TRUE),
    sd_general_ses = sd(latent_factor_ss_general_ses, na.rm = TRUE),
    mean_social = mean(latent_factor_ss_social, na.rm = TRUE),
    sd_social = sd(latent_factor_ss_social, na.rm = TRUE),
    mean_perinatal = mean(latent_factor_ss_perinatal, na.rm = TRUE),
    sd_perinatal = sd(latent_factor_ss_perinatal, na.rm = TRUE)
  )

# Print summaries
cat("Full Sample Stats for Latent Variables\n")
print(full_sample_stats_latent)

cat("\nOCD Present Stats for Latent Variables\n")
print(ocd_present_stats_latent)

cat("\nOCD Absent Stats for Latent Variables\n")
print(ocd_absent_stats_latent)
```

##	Family and Household Structure:
o	Include information on the family structure (e.g., single-parent households, guardianship) and household size if relevant.

Family environment scale
fes_p_ss_fc_pr = conflict
fes_p_ss_cohesion_sum_pr = cohesion
```{r}
fam_envi <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/culture-environment/ce_p_fes.csv")
```

## Useable MRI
```{r}
# DESKIAN ATLAS: MRI year 2 follow up
desikan_v <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/ocd/desikan_v.csv")

# Create the usableMRI variable
desikan_v <- desikan_v %>%
  group_by(src_subject_id) %>%
  mutate(usableMRI = if_else(all(!is.na(.)), 1, 0)) %>%
  ungroup()

# Select and save only src_subject_id and usableMRI
usable_mri <- desikan_v %>%
  select(src_subject_id, usableMRI)
```

```{r}
full_data <- demo %>%
  left_join(usable_mri, by = c("src_subject_id")) 
```

```{r}
# Full sample count of usable MRIs
full_sample_usable_mri <- full_data %>%
  filter(usableMRI == 1) %>%  # Only consider those with usable MRI
  summarise(total_usable_mri = n_distinct(src_subject_id)) %>%
  pull(total_usable_mri)

cat("Total number of subjects with usable MRI in the full sample:", full_sample_usable_mri, "\n")

# OCD absent sample count of usable MRIs
ocd_absent_usable_mri <- full_data %>%
  filter(ksads_11_917_p != 1, usableMRI == 1) %>%  # OCD absent and usable MRI
  summarise(total_usable_mri = n_distinct(src_subject_id)) %>%
  pull(total_usable_mri)

cat("Total number of subjects with usable MRI and OCD absent:", ocd_absent_usable_mri, "\n")

# OCD present sample count of usable MRIs
ocd_present_usable_mri <- full_data %>%
  filter(ksads_11_917_p == 1, usableMRI == 1) %>%  # OCD present and usable MRI
  summarise(total_usable_mri = n_distinct(src_subject_id)) %>%
  pull(total_usable_mri)

cat("Total number of subjects with usable MRI and OCD present:", ocd_present_usable_mri, "\n")
```

# Merge
## Demographic variables: sex, gender, ethnicity, age 
```{r}
full_data <- diagnosis %>%
  left_join(demo, by = "src_subject_id")
```

### Latent variables
1) a general latent factor of economic, social, and physiological well-being, 
2) a latent factor for youth perceived social support, and 
3) a latent factor for perinatal health.
```{r}
latent <- latent %>%
  select(-eventname)

full_data <- full_data %>%
  left_join(latent, by = c("src_subject_id")) 
```



```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```







