---
title: "Plot Externalizing Symptoms for Child-Report"
author: "MochiBear.Hei"
date: "04.16.2025"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    page-layout: full
    html-math-method: katex
editor: source
---

```{css, echo = F}
body{
  font-family: Helvetica;
  font-size: 16pt;
  max-width: 1000px;
  margin: auto;
  margin-left:310px;
}
pre{
  font-size: 14px;
}
/* Headers */
h1{
    font-size: 24pt;
  }
h1,h2{
    font-size: 20pt;
  }
h3,h4,h5,h6{
  font-size: 18pt;
}
#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 300px;
  height: 100%;
  overflow:auto;
}
```

```{r setup,suppressPackageStartupMessages= TRUE}
#| include: false
#| message: false
#| warning: false
#| results: hide
knitr::opts_chunk$set(echo = TRUE, dpi = 300)

library(caret)
library(ggplot2)
library(reshape2)
library(tidyr)
library(pROC)
library(MLmetrics)
library(dplyr)
library(tibble)
library(Metrics)
library(patchwork) 
theme_set(theme_classic(base_size = 14))
```

Evaluates the XGBoost multiclass model for BPM externalizing symptoms. Performs prediction, threshold calibration (via Youden’s J), accuracy/confusion matrix reporting, OvR ROC and calibration plots, and test set evaluation.

# load model
```{r}
load("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/outputs/models/multiclass_cbcl_bpm/m_bpm_e_model+preds.RData")
```

Make predictions
```{r}
#set labels for models = 0, 1, 2, 
class_labels <- sort(unique(train_data$bpm_externalizing))

dcalib <- xgb.DMatrix(data = x_calib, label = y_calib)

#predict on calibration set 
pred_probs_calib <- predict(final_model, dcalib, reshape = TRUE, outputmargin = FALSE)
colnames(pred_probs_calib) <- as.character(class_labels)
```

Argmax predictions
```{r}
pred_classes_argmax <- max.col(pred_probs_calib) - 1  # assumes labels 0,1,2

y_true_calib <- factor(y_calib, levels = class_labels)
y_pred_calib <- factor(pred_classes_argmax, levels = class_labels)

conf_matrix <- confusionMatrix(y_pred_calib, y_true_calib)
print(conf_matrix)
```


# Threshold calibraiton
feed softprob results (pred_probs) from final model with test data through threshold calibration
optimize balance accuracy
```{r}
# Youden's J statistic
tune_thresholds <- function(pred_probs, y_calib) {
  sapply(class_labels, function(cls) {
    # set threshold search range per class
    if (cls == "0") {
      thresholds <- seq(0.80, 1.00, by = 0.01)  # stricter range for class 0
    } else {
      thresholds <- seq(0.00, 1.00, by = 0.01)  # full range for class 1 and 2
    }

    y_true_bin <- as.integer(y_calib == cls)
    probs <- pred_probs[, as.character(cls)]
    
    youden_scores <- sapply(thresholds, function(t) {
      preds <- as.integer(probs >= t)
      
      # confusion matrix components
      tp <- sum(preds == 1 & y_true_bin == 1)
      tn <- sum(preds == 0 & y_true_bin == 0)
      fp <- sum(preds == 1 & y_true_bin == 0)
      fn <- sum(preds == 0 & y_true_bin == 1)
      
      sensitivity <- if ((tp + fn) > 0) tp / (tp + fn) else 0
      specificity <- if ((tn + fp) > 0) tn / (tn + fp) else 0
      
      # Youden's J
      youden_j <- sensitivity + specificity - 1
      return(youden_j)
    })
    
    thresholds[which.max(youden_scores)]
  })
}

best_thresholds <- tune_thresholds(pred_probs_calib, y_calib)
print(best_thresholds)
```


# Confusion matrix
```{r}
#put thresholds here
best_thresholds <- c("0" = 0.89, "1" = 0.09, "2" = 0.04)     ##################CHANGE THRESHOLDS HEEEEER

#apply threshold
thresholded_preds <- sapply(class_labels, function(cls) {
  probs <- pred_probs_calib[, as.character(cls)]
  as.integer(probs >= best_thresholds[as.character(cls)])
})

# pick highest probability
pred_classes_thresh <- apply(thresholded_preds, 1, function(row) {
  if (sum(row) == 0) {
    return(NA)  
  } else if (sum(row) == 1) {
    return(which(row == 1) - 1)  
  } else {
    return(which.max(pred_probs_calib[row == 1, ][1, ]) - 1)
  }
})

# if NA use argmax
na_indices <- is.na(pred_classes_thresh)
if (any(na_indices)) {
  fallback_preds <- max.col(pred_probs_calib[na_indices, ]) - 1
  pred_classes_thresh[na_indices] <- fallback_preds
}

y_pred_calib_thresh <- factor(pred_classes_thresh, levels = class_labels)
y_true_calib <- factor(y_calib, levels = class_labels)

conf_matrix_thresh <- confusionMatrix(y_pred_calib_thresh, y_true_calib)
print(conf_matrix_thresh)
```


# Predict probability on test set 
```{r}
dtest <- xgb.DMatrix(data = x_test, label = y_test) 

pred_probs_test <- predict(final_model, dtest, reshape = TRUE)
colnames(pred_probs_test) <- as.character(class_labels)

thresholded_preds_test <- sapply(class_labels, function(cls) {
  probs <- pred_probs_test[, as.character(cls)]
  as.integer(probs >= best_thresholds[as.character(cls)])
})

pred_classes_test <- apply(thresholded_preds_test, 1, function(row) {
  if (sum(row) == 0) {
    return(NA)  
  } else if (sum(row) == 1) {
    return(which(row == 1) - 1)
  } else {
    return(which.max(pred_probs_test[row == 1, ][1, ]) - 1)
  }
})

na_indices <- is.na(pred_classes_test)
if (any(na_indices)) {
  fallback_preds <- max.col(pred_probs_test[na_indices, ]) - 1
  pred_classes_test[na_indices] <- fallback_preds
}
```

# Evaluate
```{r}
y_true_test <- factor(y_test, levels = class_labels)
y_pred_test <- factor(pred_classes_test, levels = class_labels)

conf_matrix_test <- confusionMatrix(y_pred_test, y_true_test)
print(conf_matrix_test)
```



for plotting
```{r}
df_plot <- as.data.frame(pred_probs_test)
colnames(df_plot) <- as.character(class_labels)
df_plot$true_label <- y_test
# long format
df_long <- df_plot %>%
  pivot_longer(cols = -true_label, names_to = "class", values_to = "prob") %>%
  mutate(class = factor(class, levels = class_labels),
         is_true_class = as.integer(as.character(true_label) == as.character(class)))
```

# ROC AUC 
```{r}
#plot all together 
class_names <- c("0" = "Healthy", "1" = "Borderline", "2" = "Clinical")
plots <- lapply(class_labels, function(cls) {
  df_cls <- df_long %>% filter(class == cls)
  cls_label <- class_names[as.character(cls)]
  
  # Histogram
  hist_plot <- ggplot(df_cls, aes(x = prob, fill = factor(is_true_class))) +
    geom_histogram(position = "identity", alpha = 0.8, bins = 30, color = "black") +
    scale_fill_manual(
      values = c("0" = "red", "1" = "#800080"),
      name = "",
      labels = c("Rest", paste("Class:", cls_label))
    ) +
    labs(
      x = paste0("P(x = ", cls_label, ")"),
      y = "Count",
      title = cls_label
    ) +
    theme_classic(base_size = 13) +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = "top"
    )
  
  # ROC
  roc_obj <- roc(df_cls$is_true_class, df_cls$prob, quiet = TRUE)
  roc_df <- data.frame(
    FPR = 1 - roc_obj$specificities,
    TPR = roc_obj$sensitivities
  )
  roc_plot <- ggplot(roc_df, aes(x = FPR, y = TPR)) +
    geom_line(color = "blue", size = 1) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "darkgreen") +
    labs(
      x = "False Positive Rate",
      y = "True Positive Rate",
      title = "ROC Curve OvR"
    ) +
    theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(hjust = 0.5)
    )
  
  # Combine histogram + ROC
  hist_plot / roc_plot
})

final_plot <- wrap_plots(plots, nrow = 1)
print(final_plot)

ggsave("bpm_e_OvR_ROC.png", final_plot, width = 12, height = 6, dpi = 300)

```



#plot confusion matrix
```{r}
conf_mat_table <- table(y_true_test, y_pred_test)

conf_df <- as.data.frame(conf_mat_table)
colnames(conf_df) <- c("True", "Predicted", "Count")

class_names <- c("0" = "Healthy", "1" = "Borderline", "2" = "Clinical")
conf_df$True <- factor(conf_df$True, levels = names(class_names), labels = class_names)
conf_df$Predicted <- factor(conf_df$Predicted, levels = names(class_names), labels = class_names)
ggplot(conf_df, aes(x = Predicted, y = True, fill = Count)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Count), size = 5) +
  scale_fill_gradient(low = "white", high = "#9966CC") +
  labs(
    title = "Confusion Matrix (Test Set)",
    x = "Predicted Class",
    y = "True Class",
    fill = "Count"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

ggsave("confusion_matrix_heatmap.png", width = 6, height = 5, dpi = 300)

```