---
title: "Plot Internalizing Symptoms for Children-Report"
author: "MochiBear.Hei"
date: "04.16.2025"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    page-layout: full
    html-math-method: katex
editor: source
---

```{css, echo = F}
body{
  font-family: Helvetica;
  font-size: 16pt;
  max-width: 1000px;
  margin: auto;
  margin-left:310px;
}
pre{
  font-size: 14px;
}
/* Headers */
h1{
    font-size: 24pt;
  }
h1,h2{
    font-size: 20pt;
  }
h3,h4,h5,h6{
  font-size: 18pt;
}
#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 300px;
  height: 100%;
  overflow:auto;
}
```

```{r setup,suppressPackageStartupMessages= TRUE}
#| include: false
#| message: false
#| warning: false
#| results: hide
knitr::opts_chunk$set(echo = TRUE, dpi = 300)

library(caret)
library(ggplot2)
library(reshape2)
library(tidyr)
library(pROC)
library(MLmetrics)
library(dplyr)
library(tibble)
library(Metrics)
library(patchwork) 
theme_set(theme_classic(base_size = 14))
```

Evaluates XGBoost multiclass model for BPM internalizing symptoms. Includes argmax and threshold-calibrated prediction accuracy, confusion matrices, ROC curves, calibration curves, and a permutation test for statistical significance.

# load model
```{r}
load("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/outputs/models/multiclass_cbcl_bpm/m_bpm_i_model+preds.RData")
```

Make predictions
```{r}
#set labels for models = 0, 1, 2, 
class_labels <- sort(unique(train_data$bpm_internalizing))

#xgboost test matrix 
dcalib <- xgb.DMatrix(data = x_calib, label = y_calib)

# ---- Predict probabilities on calibration set ----
pred_probs_calib <- predict(final_model, dcalib, reshape = TRUE, outputmargin = FALSE)
colnames(pred_probs_calib) <- as.character(class_labels)
```

# Argmax predictions
```{r}
# Argmax predictions (raw, not calibrated)
pred_classes_argmax <- max.col(pred_probs_calib) - 1  # assumes labels 0,1,2

# Factors for comparison
y_true_calib <- factor(y_calib, levels = class_labels)
y_pred_calib <- factor(pred_classes_argmax, levels = class_labels)

# Accuracy and confusion matrix
accuracy <- mean(y_pred_calib == y_true_calib)
cat("✅ Accuracy:", round(accuracy, 4), "\n")

conf_matrix <- confusionMatrix(y_pred_calib, y_true_calib)
print(conf_matrix)
```

# Threshold calibraiton
feed softprob results (pred_probs) from final model with test data through threshold calibration
optimize balanced accuracy 
```{r}
# thresholds per class using Youden's J statistic
tune_thresholds <- function(pred_probs, y_calib) {
  sapply(class_labels, function(cls) {
    # Set threshold search range per class
    if (cls == "0") {
      thresholds <- seq(0.80, 1.00, by = 0.01)  # stricter range for class 0
    } else {
      thresholds <- seq(0.00, 1.00, by = 0.01)  # full range for class 1 and 2
    }

    y_true_bin <- as.integer(y_calib == cls)
    probs <- pred_probs[, as.character(cls)]
    
    youden_scores <- sapply(thresholds, function(t) {
      preds <- as.integer(probs >= t)
      
      # Confusion matrix components
      tp <- sum(preds == 1 & y_true_bin == 1)
      tn <- sum(preds == 0 & y_true_bin == 0)
      fp <- sum(preds == 1 & y_true_bin == 0)
      fn <- sum(preds == 0 & y_true_bin == 1)
      
      sensitivity <- if ((tp + fn) > 0) tp / (tp + fn) else 0
      specificity <- if ((tn + fp) > 0) tn / (tn + fp) else 0
      
      # Youden's J
      youden_j <- sensitivity + specificity - 1
      return(youden_j)
    })
    
    thresholds[which.max(youden_scores)]
  })
}


# Calibrate thresholds
best_thresholds <- tune_thresholds(pred_probs_calib, y_calib)
print(best_thresholds)

```

# calibration confusion matrix
```{r}

# manual thresholds (corresponding to class labels 0, 1, 2)
best_thresholds <- c("0" = 0.83, "1" = 0.02, "2" = 0.02)     ##################CHANGE THRESHOLDS HEEEEER
#apply
thresholded_preds <- sapply(class_labels, function(cls) {
  probs <- pred_probs_calib[, as.character(cls)]
  as.integer(probs >= best_thresholds[as.character(cls)])
})

pred_classes_thresh <- apply(thresholded_preds, 1, function(row) {
  if (sum(row) == 0) {
    return(NA)  # No threshold passed
  } else if (sum(row) == 1) {
    return(which(row == 1) - 1)  # single class passed threshold
  } else {
    # Multiple thresholds passed: pick one with max probability
    return(which.max(pred_probs_calib[row == 1, ][1, ]) - 1)
  }
})

# if NA use argmax 
na_indices <- is.na(pred_classes_thresh)
if (any(na_indices)) {
  fallback_preds <- max.col(pred_probs_calib[na_indices, ]) - 1
  pred_classes_thresh[na_indices] <- fallback_preds
}

y_pred_calib_thresh <- factor(pred_classes_thresh, levels = class_labels)
y_true_calib <- factor(y_calib, levels = class_labels)

# evaluate
conf_matrix_thresh <- confusionMatrix(y_pred_calib_thresh, y_true_calib)
print(conf_matrix_thresh)
```

# Predict probability on test set 
```{r}
dtest <- xgb.DMatrix(data = x_test, label = y_test) 

# predict soft probabilities
pred_probs_test <- predict(final_model, dtest, reshape = TRUE)
colnames(pred_probs_test) <- as.character(class_labels)

thresholded_preds_test <- sapply(class_labels, function(cls) {
  probs <- pred_probs_test[, as.character(cls)]
  as.integer(probs >= best_thresholds[as.character(cls)])
})

pred_classes_test <- apply(thresholded_preds_test, 1, function(row) {
  if (sum(row) == 0) {
    return(NA)  # no threshold met
  } else if (sum(row) == 1) {
    return(which(row == 1) - 1)
  } else {
    return(which.max(pred_probs_test[row == 1, ][1, ]) - 1)
  }
})

na_indices <- is.na(pred_classes_test)
if (any(na_indices)) {
  fallback_preds <- max.col(pred_probs_test[na_indices, ]) - 1
  pred_classes_test[na_indices] <- fallback_preds
}


```

# Evaluate
```{r}
y_true_test <- factor(y_test, levels = class_labels)
y_pred_test <- factor(pred_classes_test, levels = class_labels)

conf_matrix_test <- confusionMatrix(y_pred_test, y_true_test)
print(conf_matrix_test)
```

for plotting
```{r}
df_plot <- as.data.frame(pred_probs_test)
colnames(df_plot) <- as.character(class_labels)
df_plot$true_label <- y_test

# long format
df_long <- df_plot %>%
  pivot_longer(cols = -true_label, names_to = "class", values_to = "prob") %>%
  mutate(class = factor(class, levels = class_labels),
         is_true_class = as.integer(as.character(true_label) == as.character(class)))

```

#plot all together 
```{r}
class_names <- c("0" = "Healthy", "1" = "Borderline", "2" = "Clinical")
plots <- lapply(class_labels, function(cls) {
  df_cls <- df_long %>% filter(class == cls)
  cls_label <- class_names[as.character(cls)]
  
  # Histogram
  hist_plot <- ggplot(df_cls, aes(x = prob, fill = factor(is_true_class))) +
    geom_histogram(position = "identity", alpha = 0.8, bins = 30, color = "black") +
    scale_fill_manual(
      values = c("0" = "red", "1" = "#800080"),
      name = "",
      labels = c("Rest", paste("Class:", cls_label))
    ) +
    labs(
      x = paste0("P(x = ", cls_label, ")"),
      y = "Count",
      title = cls_label
    ) +
    theme_classic(base_size = 13) +
    theme(
      plot.title = element_text(hjust = 0.5),
      legend.position = "top"
    )
  
  # ROC
  roc_obj <- roc(df_cls$is_true_class, df_cls$prob, quiet = TRUE)
  roc_df <- data.frame(
    FPR = 1 - roc_obj$specificities,
    TPR = roc_obj$sensitivities
  )
  roc_plot <- ggplot(roc_df, aes(x = FPR, y = TPR)) +
    geom_line(color = "blue", size = 1) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "darkgreen") +
    labs(
      x = "False Positive Rate",
      y = "True Positive Rate",
      title = "ROC Curve OvR"
    ) +
    theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(hjust = 0.5)
    )
  
  # Combine histogram + ROC
  hist_plot / roc_plot
})

final_plot <- wrap_plots(plots, nrow = 1)
print(final_plot)

ggsave("bpm_i_OvR_ROC.png", final_plot, width = 12, height = 6, dpi = 300)

```

#plot confusion matrix
```{r}
conf_mat_table <- table(y_true_test, y_pred_test)

conf_df <- as.data.frame(conf_mat_table)
colnames(conf_df) <- c("True", "Predicted", "Count")

class_names <- c("0" = "Healthy", "1" = "Borderline", "2" = "Clinical")
conf_df$True <- factor(conf_df$True, levels = names(class_names), labels = class_names)
conf_df$Predicted <- factor(conf_df$Predicted, levels = names(class_names), labels = class_names)
ggplot(conf_df, aes(x = Predicted, y = True, fill = Count)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Count), size = 5) +
  scale_fill_gradient(low = "white", high = "#9966CC") +
  labs(
    title = "Confusion Matrix (Test Set)",
    x = "Predicted Class",
    y = "True Class",
    fill = "Count"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

ggsave("confusion_matrix_heatmap.png", width = 6, height = 5, dpi = 300)

```

#plot calibration curves
```{r}
get_calibration_data <- function(df, class_label, n_bins = 10) {
  df_class <- df %>%
    filter(class == class_label) %>%
    mutate(bin = cut(prob, breaks = seq(0, 1, length.out = n_bins + 1), include.lowest = TRUE)) %>%
    group_by(bin) %>%
    summarise(
      mean_pred = mean(prob),
      frac_positives = mean(is_true_class),
      .groups = "drop"
    )
  
  df_class$class <- class_label
  return(df_class)
}

# Apply to each class
cal_data <- bind_rows(lapply(class_labels, function(cls) get_calibration_data(df_long, cls)))
class_names <- c("0" = "Healthy", "1" = "Borderline", "2" = "Clinical")

cal_data$class <- factor(cal_data$class, levels = names(class_names), labels = class_names)

ggplot(cal_data, aes(x = mean_pred, y = frac_positives, color = class)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "black") +
  labs(
    title = "Calibration Curves (OvR)",
    x = "Mean Predicted Probability",
    y = "Fraction of Positives (Observed)",
    color = "Class"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "bottom"
  )

ggsave("calibration_curves_OvR.png", width = 7, height = 5, dpi = 300)

```


# Permutation test 
```{r}
# Observed accuracy
observed_acc <- mean(pred_classes_test == y_test)

# Permutation loop
n_perms <- 1000
perm_accs <- numeric(n_perms)

for (i in 1:n_perms) {
  permuted_labels <- sample(y_test)
  dtest_perm <- xgb.DMatrix(data = x_test, label = permuted_labels)
  
  pred_probs_perm <- predict(final_model, dtest_perm, reshape = TRUE)
  colnames(pred_probs_perm) <- as.character(class_labels)

  thresholded_preds_perm <- sapply(class_labels, function(cls) {
    probs <- pred_probs_perm[, as.character(cls)]
    as.integer(probs >= best_thresholds[as.character(cls)])
  })

  pred_classes_perm <- apply(thresholded_preds_perm, 1, function(row) {
    if (sum(row) == 0) return(NA)
    else if (sum(row) == 1) return(which(row == 1) - 1)
    else return(which.max(pred_probs_perm[row == 1, ][1, ]) - 1)
  })

  na_idx <- is.na(pred_classes_perm)
  if (any(na_idx)) {
    fallback_preds <- max.col(pred_probs_perm[na_idx, ]) - 1
    pred_classes_perm[na_idx] <- fallback_preds
  }

  perm_accs[i] <- mean(pred_classes_perm == permuted_labels)
}

# Compute p-value
p_value <- mean(perm_accs >= observed_acc)
print(paste("Permutation test p-value:", p_value))
```

plot perumation
```{r}
hist_data <- data.frame(stat = perm_accs)

dens <- density(perm_accs)
y_pos <- max(dens$y) * 0.95 

perm_plot <- ggplot(hist_data, aes(x = stat)) +
  geom_density(fill = "#800080", alpha = 0.4) +
  geom_vline(xintercept = observed_acc, color = "red", size = 1.2) +
  annotate("text", x = observed_acc, y = y_pos,
           label = "Observed accuracy", hjust = -0.1, vjust = 1, color = "red", size = 4) +
  labs(
    title = "Permutation Test: Model Accuracy vs. Null Distribution",
    subtitle = "Null distribution of accuracy under label permutation",
    x = "Classification Accuracy",
    y = "Density",
    caption = "Green line: Accuracy on true labels\nBlue area: Distribution under random label permutations"
  ) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12),
    plot.caption = element_text(size = 10, hjust = 0, color = "gray30")
  )

print(perm_plot)

ggsave("permutation_test.png", plot = perm_plot, width = 7, height = 5, dpi = 300)
```

