


---
title: "Demographic characteristics"
author: "MochiBear.Hei"
date: "12.20.2024"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    page-layout: full
    html-math-method: katex
editor: source
---

```{css, echo = F}
body{
  font-family: Helvetica;
  font-size: 16pt;
  max-width: 1000px;
  margin: auto;
  margin-left:310px;
}
pre{
  font-size: 14px;
}
/* Headers */
h1{
    font-size: 24pt;
  }
h1,h2{
    font-size: 20pt;
  }
h3,h4,h5,h6{
  font-size: 18pt;
}
#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 300px;
  height: 100%;
  overflow:auto;
}
```


```{r setup,suppressPackageStartupMessages= TRUE}
#| include: false
#| message: false
#| warning: false
#| results: hide
knitr::opts_chunk$set(echo = TRUE, dpi = 300)

# Data manipulation
library(dplyr)
library(tidyr)

# Plotting
library(ggplot2)

# Utilities
library(knitr)

library(gt)
library(epitools)
```

#
Analyzes psychiatric comorbidities in OCD vs. non-OCD groups using summary diagnosis flags and CBCL DSM scales. Computes prevalence, t-tests, and group differences across multiple disorders.

## Load
```{r}

all_ksads_diagnosis <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/comorbidities.csv") %>%
  select(
    src_subject_id,
    ksads_11_917_p,
    adhd,
    any_bipolar,
    any_conduct_odd,
    any_depressive_disorder,
    any_anxiety_disorder,
    any_drug_use_disorder,
    any_eating_disorder,
    ptsd,
    any_suicidality,
    no_diagnosis,
    only_ocd,
    comorbidities
  ) %>%
  filter(!is.na(ksads_11_917_p))

##list of people to include from flow script
subids <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/subids.csv") %>%
  select(src_subject_id) 

all_ksads_diagnosis <- inner_join(subids, all_ksads_diagnosis, by = "src_subject_id")

## how many with no ocd
total_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1) %>% # Filter for cases where OCD is absent
  summarise(total_ocd_absent = n_distinct(src_subject_id)) %>%
  pull(total_ocd_absent)

cat("Total number of people with OCD absent:", total_ocd_absent, "\n")


```

##Check individually
Depressive disorders 
```{r}
# Total number of people in the full sample with any depressive disorder
total_depressive_disorder_full <- all_ksads_diagnosis %>%
  filter(any_depressive_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any depressive disorder
total_depressive_disorder_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, any_depressive_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any depressive disorder
total_depressive_disorder_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, any_depressive_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with any depressive disorder:", total_depressive_disorder_full, "\n")
cat("Total number of people with OCD and any depressive disorder:", total_depressive_disorder_ocd_present, "\n")
cat("Total number of people without OCD and with any depressive disorder:", total_depressive_disorder_ocd_absent, "\n")

```

Anxiety disorders
```{r}
# Total number of people in the full sample with any anxiety disorder
total_anxiety_disorder_full <- all_ksads_diagnosis %>%
  filter(any_anxiety_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any anxiety disorder
total_anxiety_disorder_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, any_anxiety_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any anxiety disorder
total_anxiety_disorder_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, any_anxiety_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with any anxiety disorder:", total_anxiety_disorder_full, "\n")
cat("Total number of people with OCD and any anxiety disorder:", total_anxiety_disorder_ocd_present, "\n")
cat("Total number of people without OCD and with any anxiety disorder:", total_anxiety_disorder_ocd_absent, "\n")

```

ADHD
```{r}
# Total number of people in the full sample with any anxiety disorder
total_adhd_full <- all_ksads_diagnosis %>%
  filter(adhd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any anxiety disorder
total_adhd_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, adhd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any anxiety disorder
total_adhd_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, adhd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with adhd:", total_adhd_full, "\n")
cat("Total number of people with OCD and adhd:", total_adhd_ocd_present, "\n")
cat("Total number of people without OCD and adhd:", total_adhd_ocd_absent, "\n")
```

ODD /CD
any_conduct_odd
```{r}
# Total number of people in the full sample with any anxiety disorder
total_odd_full <- all_ksads_diagnosis %>%
  filter(any_conduct_odd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any anxiety disorder
total_odd_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, any_conduct_odd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any anxiety disorder
total_odd_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, any_conduct_odd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with odd:", total_odd_full, "\n")
cat("Total number of people with OCD and odd:", total_odd_ocd_present, "\n")
cat("Total number of people without OCD and odd:", total_odd_ocd_absent, "\n")
```

PTSD
ptsd
```{r}
# Total number of people in the full sample with any anxiety disorder
total_ptsd_full <- all_ksads_diagnosis %>%
  filter(ptsd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any anxiety disorder
total_ptsd_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, ptsd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any anxiety disorder
total_ptsd_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, ptsd == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with ptsd:", total_ptsd_full, "\n")
cat("Total number of people with OCD and ptsd:", total_ptsd_ocd_present, "\n")
cat("Total number of people without OCD and ptsd:", total_ptsd_ocd_absent, "\n")
```

Drug use disorder
"any_drug_use_disorder" 
```{r}
# Total number of people in the full sample with any anxiety disorder
total_any_drug_use_disorder_full <- all_ksads_diagnosis %>%
  filter(any_drug_use_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any anxiety disorder
total_any_drug_use_disorder_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, any_drug_use_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any anxiety disorder
total_any_drug_use_disorder_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, any_drug_use_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with any_drug_use_disorder:", total_any_drug_use_disorder_full, "\n")
cat("Total number of people with OCD and any_drug_use_disorder:", total_any_drug_use_disorder_ocd_present, "\n")
cat("Total number of people without OCD and any_drug_use_disorder:", total_any_drug_use_disorder_ocd_absent, "\n")
```

Bipolar
"any_bipolar" 
```{r}
# Total number of people in the full sample with any anxiety disorder
total_any_bipolar_full <- all_ksads_diagnosis %>%
  filter(any_bipolar == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any anxiety disorder
total_any_bipolar_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, any_bipolar == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any anxiety disorder
total_any_bipolar_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, any_bipolar == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with any_bipolar:", total_any_bipolar_full, "\n")
cat("Total number of people with OCD and any_bipolar:", total_any_bipolar_ocd_present, "\n")
cat("Total number of people without OCD and any_bipolar:", total_any_bipolar_ocd_absent, "\n")
```

Eating disorder
"any_eating_disorder"
```{r}
# Total number of people in the full sample with any anxiety disorder
total_any_eating_disorder_full <- all_ksads_diagnosis %>%
  filter(any_eating_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any anxiety disorder
total_any_eating_disorder_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, any_eating_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any anxiety disorder
total_any_eating_disorder_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, any_eating_disorder == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with any_eating_disorder:", total_any_eating_disorder_full, "\n")
cat("Total number of people with OCD and any_eating_disorder:", total_any_eating_disorder_ocd_present, "\n")
cat("Total number of people without OCD and any_eating_disorder:", total_any_eating_disorder_ocd_absent, "\n")
```

Suicidality
"any_suicidality"  
```{r}
# Total number of people in the full sample with any anxiety disorder
total_any_suicidality_full <- all_ksads_diagnosis %>%
  filter(any_suicidality == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with any anxiety disorder
total_any_suicidality_ocd_present <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p == 1, any_suicidality == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD absent sample with any anxiety disorder
total_any_suicidality_ocd_absent <- all_ksads_diagnosis %>%
  filter(ksads_11_917_p != 1, any_suicidality == 1) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Display the results
cat("Total number of people in the full sample with any_suicidality:", total_any_suicidality_full, "\n")
cat("Total number of people with OCD and any_suicidality:", total_any_suicidality_ocd_present, "\n")
cat("Total number of people without OCD and any_suicidality:", total_any_suicidality_ocd_absent, "\n")
```
No diagnosis
```{r}
# Defining a helper function to calculate totals based on specified conditions
calculate_total <- function(data, ocd_condition) {
  data %>%
    filter(
      ksads_11_917_p == ocd_condition,
      adhd == 0,
      any_bipolar == 0,
      any_conduct_odd == 0,
      any_depressive_disorder == 0,
      any_anxiety_disorder == 0,
      any_drug_use_disorder == 0,
      any_eating_disorder == 0,
      ptsd == 0,
      any_suicidality == 0
    ) %>%
    summarise(total = n_distinct(src_subject_id)) %>%
    pull(total)
}

# Total number of people in the full sample with no diagnosis
total_no_diagnosis_full <- all_ksads_diagnosis %>%
  filter(
    adhd == 0,
    any_bipolar == 0,
    any_conduct_odd == 0,
    any_depressive_disorder == 0,
    any_anxiety_disorder == 0,
    any_drug_use_disorder == 0,
    any_eating_disorder == 0,
    ptsd == 0,
    any_suicidality == 0
  ) %>%
  summarise(total = n_distinct(src_subject_id)) %>%
  pull(total)

# Total number of people in the OCD present sample with no other diagnosis
total_no_other_diagnosis_ocd_present <- calculate_total(all_ksads_diagnosis, 1)

# Total number of people in the OCD absent sample with no diagnosis
total_no_diagnosis_ocd_absent <- calculate_total(all_ksads_diagnosis, 0)

# Display the results
cat("Total number of people in the full sample with no diagnosis:", total_no_diagnosis_full, "\n")
cat("Total number of people with OCD and no other diagnosis:", total_no_other_diagnosis_ocd_present, "\n")
cat("Total number of people without OCD and no diagnosis:", total_no_diagnosis_ocd_absent, "\n")
```

# CBCL
```{r}
dsm_diagnosis <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/core/mental-health/mh_p_cbcl.csv") %>%
  filter(eventname == "2_year_follow_up_y_arm_1") %>%
  select(src_subject_id, cbcl_scr_syn_attention_t, cbcl_scr_dsm5_anxdisord_t, cbcl_scr_dsm5_adhd_t, cbcl_scr_dsm5_depress_t, cbcl_scr_dsm5_somaticpr_t, cbcl_scr_dsm5_opposit_t, cbcl_scr_dsm5_conduct_t, cbcl_scr_syn_internal_t, cbcl_scr_syn_external_t, cbcl_scr_syn_totprob_t, cbcl_scr_07_ocd_t)

##list of people to include from flow script
subids <- read.csv("/Users/maggieheimvik/Desktop/GitHub/masters_thesis/Data/ocd/subids.csv") %>%
  select(src_subject_id, ksads_11_917_p) 


dsm_diagnosis <- inner_join(subids, dsm_diagnosis, by = "src_subject_id")
```

Merge DSM diagnosis with OCD diagnosis
```{r}
## how many with no ocd
# Calculate the total number of people with OCD absent
total_ocd_absent <- dsm_diagnosis %>%
  filter(ksads_11_917_p != 1) %>% # Filter for cases where OCD is absent
  summarise(total_ocd_absent = n_distinct(src_subject_id)) %>%
  pull(total_ocd_absent)

# Display the result
cat("Total number of people with OCD absent:", total_ocd_absent, "\n")

## how many with ocd
# Calculate the total number of people with OCD absent
total_ocd_present <- dsm_diagnosis %>%
  filter(ksads_11_917_p == 1) %>% # Filter for cases where OCD is absent
  summarise(total_ocd_present = n_distinct(src_subject_id)) %>%
  pull(total_ocd_present)

# Display the result
cat("Total number of people with OCD present:", total_ocd_present, "\n")


```

## DSM comorbidity
```{r}

# Full dataset statistics for each T score category
full_sample_stats <- dsm_diagnosis %>%
  summarise(
    mean_attention_t = mean(cbcl_scr_syn_attention_t, na.rm = TRUE),
    sd_attention_t = sd(cbcl_scr_syn_attention_t, na.rm = TRUE),
    mean_anxdisord_t = mean(cbcl_scr_dsm5_anxdisord_t, na.rm = TRUE),
    sd_anxdisord_t = sd(cbcl_scr_dsm5_anxdisord_t, na.rm = TRUE),
    mean_adhd_t = mean(cbcl_scr_dsm5_adhd_t, na.rm = TRUE),
    sd_adhd_t = sd(cbcl_scr_dsm5_adhd_t, na.rm = TRUE),
    mean_depress_t = mean(cbcl_scr_dsm5_depress_t, na.rm = TRUE),
    sd_depress_t = sd(cbcl_scr_dsm5_depress_t, na.rm = TRUE),
    mean_somaticpr_t = mean(cbcl_scr_dsm5_somaticpr_t, na.rm = TRUE),
    sd_somaticpr_t = sd(cbcl_scr_dsm5_somaticpr_t, na.rm = TRUE),
    mean_opposit_t = mean(cbcl_scr_dsm5_opposit_t, na.rm = TRUE),
    sd_opposit_t = sd(cbcl_scr_dsm5_opposit_t, na.rm = TRUE),
    mean_conduct_t = mean(cbcl_scr_dsm5_conduct_t, na.rm = TRUE),
    sd_conduct_t = sd(cbcl_scr_dsm5_conduct_t, na.rm = TRUE),
    mean_internal_t = mean(cbcl_scr_syn_internal_t, na.rm = TRUE),
    sd_internal_t = sd(cbcl_scr_syn_internal_t, na.rm = TRUE),
    mean_external_t = mean(cbcl_scr_syn_external_t, na.rm = TRUE),
    sd_external_t = sd(cbcl_scr_syn_external_t, na.rm = TRUE),
    mean_totprob_t = mean(cbcl_scr_syn_totprob_t, na.rm = TRUE),
    sd_totprob_t = sd(cbcl_scr_syn_totprob_t, na.rm = TRUE),
    mean_ocd_t = mean(cbcl_scr_07_ocd_t, na.rm = TRUE),
    sd_ocd_t = sd(cbcl_scr_07_ocd_t, na.rm = TRUE)
  )

# Statistics for OCD present population
ocd_present_stats <- dsm_diagnosis %>%
  filter(ksads_11_917_p == 1) %>%
  summarise(
    mean_attention_t = mean(cbcl_scr_syn_attention_t, na.rm = TRUE),
    sd_attention_t = sd(cbcl_scr_syn_attention_t, na.rm = TRUE),
    mean_anxdisord_t = mean(cbcl_scr_dsm5_anxdisord_t, na.rm = TRUE),
    sd_anxdisord_t = sd(cbcl_scr_dsm5_anxdisord_t, na.rm = TRUE),
    mean_adhd_t = mean(cbcl_scr_dsm5_adhd_t, na.rm = TRUE),
    sd_adhd_t = sd(cbcl_scr_dsm5_adhd_t, na.rm = TRUE),
    mean_depress_t = mean(cbcl_scr_dsm5_depress_t, na.rm = TRUE),
    sd_depress_t = sd(cbcl_scr_dsm5_depress_t, na.rm = TRUE),
    mean_somaticpr_t = mean(cbcl_scr_dsm5_somaticpr_t, na.rm = TRUE),
    sd_somaticpr_t = sd(cbcl_scr_dsm5_somaticpr_t, na.rm = TRUE),
    mean_opposit_t = mean(cbcl_scr_dsm5_opposit_t, na.rm = TRUE),
    sd_opposit_t = sd(cbcl_scr_dsm5_opposit_t, na.rm = TRUE),
    mean_conduct_t = mean(cbcl_scr_dsm5_conduct_t, na.rm = TRUE),
    sd_conduct_t = sd(cbcl_scr_dsm5_conduct_t, na.rm = TRUE),
    mean_internal_t = mean(cbcl_scr_syn_internal_t, na.rm = TRUE),
    sd_internal_t = sd(cbcl_scr_syn_internal_t, na.rm = TRUE),
    mean_external_t = mean(cbcl_scr_syn_external_t, na.rm = TRUE),
    sd_external_t = sd(cbcl_scr_syn_external_t, na.rm = TRUE),
    mean_totprob_t = mean(cbcl_scr_syn_totprob_t, na.rm = TRUE),
    sd_totprob_t = sd(cbcl_scr_syn_totprob_t, na.rm = TRUE),
    mean_ocd_t = mean(cbcl_scr_07_ocd_t, na.rm = TRUE),
    sd_ocd_t = sd(cbcl_scr_07_ocd_t, na.rm = TRUE)
  )

# Statistics for OCD absent population
ocd_absent_stats <- dsm_diagnosis %>%
  filter(ksads_11_917_p == 0) %>%
  summarise(
    mean_attention_t = mean(cbcl_scr_syn_attention_t, na.rm = TRUE),
    sd_attention_t = sd(cbcl_scr_syn_attention_t, na.rm = TRUE),
    mean_anxdisord_t = mean(cbcl_scr_dsm5_anxdisord_t, na.rm = TRUE),
    sd_anxdisord_t = sd(cbcl_scr_dsm5_anxdisord_t, na.rm = TRUE),
    mean_adhd_t = mean(cbcl_scr_dsm5_adhd_t, na.rm = TRUE),
    sd_adhd_t = sd(cbcl_scr_dsm5_adhd_t, na.rm = TRUE),
    mean_depress_t = mean(cbcl_scr_dsm5_depress_t, na.rm = TRUE),
    sd_depress_t = sd(cbcl_scr_dsm5_depress_t, na.rm = TRUE),
    mean_somaticpr_t = mean(cbcl_scr_dsm5_somaticpr_t, na.rm = TRUE),
    sd_somaticpr_t = sd(cbcl_scr_dsm5_somaticpr_t, na.rm = TRUE),
    mean_opposit_t = mean(cbcl_scr_dsm5_opposit_t, na.rm = TRUE),
    sd_opposit_t = sd(cbcl_scr_dsm5_opposit_t, na.rm = TRUE),
    mean_conduct_t = mean(cbcl_scr_dsm5_conduct_t, na.rm = TRUE),
    sd_conduct_t = sd(cbcl_scr_dsm5_conduct_t, na.rm = TRUE),
    mean_internal_t = mean(cbcl_scr_syn_internal_t, na.rm = TRUE),
    sd_internal_t = sd(cbcl_scr_syn_internal_t, na.rm = TRUE),
    mean_external_t = mean(cbcl_scr_syn_external_t, na.rm = TRUE),
    sd_external_t = sd(cbcl_scr_syn_external_t, na.rm = TRUE),
    mean_totprob_t = mean(cbcl_scr_syn_totprob_t, na.rm = TRUE),
    sd_totprob_t = sd(cbcl_scr_syn_totprob_t, na.rm = TRUE),
    mean_ocd_t = mean(cbcl_scr_07_ocd_t, na.rm = TRUE),
    sd_ocd_t = sd(cbcl_scr_07_ocd_t, na.rm = TRUE)
  )

# Print summaries
print("Full Sample Statistics")
print(full_sample_stats)

print("OCD Present Statistics")
print(ocd_present_stats)

print("OCD Absent Statistics")
print(ocd_absent_stats)

```
### t-test 
```{r}
# Perform t-test for each CBCL diagnostic category
t_test_results_cbcl <- dsm_diagnosis %>%
  mutate(has_ocd = if_else(ksads_11_917_p == 1, "OCD", "No OCD")) %>%
  pivot_longer(
    cols = c(cbcl_scr_dsm5_anxdisord_t, cbcl_scr_dsm5_adhd_t, cbcl_scr_dsm5_depress_t, 
             cbcl_scr_dsm5_somaticpr_t, cbcl_scr_dsm5_opposit_t, cbcl_scr_dsm5_conduct_t),
    names_to = "comorbidity",
    values_to = "score"
  ) %>%
  group_by(comorbidity) %>%
  summarise(
    t_test_result = list(t.test(score ~ has_ocd, data = .)),
    .groups = "drop"
  ) %>%
  mutate(
    p_value = map_dbl(t_test_result, ~ .x$p.value),
    t_statistic = map_dbl(t_test_result, ~ .x$statistic)
  ) %>%
  select(comorbidity, p_value, t_statistic)

# View results
print(t_test_results_cbcl)
```

### Cohens D (effect size)
```{r}

```


```{r}
# Create a function to summarize diagnoses
summarize_diagnosis <- function(data, diagnosis_col, group_col) {
  data %>%
    group_by(!!sym(group_col)) %>%
    summarise(
      n = sum(!is.na(!!sym(diagnosis_col)) & !!sym(diagnosis_col) == 1, na.rm = TRUE),
      pct = n / nrow(data) * 100
    )
}

# Example: Summarize any depressive disorder (ksads_1_840_p) by OCD presence
depressive_summary <- summarize_diagnosis(merged_diagnosis, "ksads_1_840_p", "ksads_11_917_p")


# Perform t-tests and extract p-values and statistics
cbcl_t_tests <- merged_diagnosis %>%
  summarise(
    Anxiety_mean_diff = mean(cbcl_scr_dsm5_anxdisord_t[ksads_11_917_p == 1], na.rm = TRUE) - 
                        mean(cbcl_scr_dsm5_anxdisord_t[ksads_11_917_p == 0], na.rm = TRUE),
    Depression_mean_diff = mean(cbcl_scr_dsm5_depress_t[ksads_11_917_p == 1], na.rm = TRUE) - 
                           mean(cbcl_scr_dsm5_depress_t[ksads_11_917_p == 0], na.rm = TRUE),
    Anxiety_p_value = t.test(cbcl_scr_dsm5_anxdisord_t ~ ksads_11_917_p, data = ., na.action = na.omit)$p.value,
    Depression_p_value = t.test(cbcl_scr_dsm5_depress_t ~ ksads_11_917_p, data = ., na.action = na.omit)$p.value,
    Anxiety_t_stat = t.test(cbcl_scr_dsm5_anxdisord_t ~ ksads_11_917_p, data = ., na.action = na.omit)$statistic,
    Depression_t_stat = t.test(cbcl_scr_dsm5_depress_t ~ ksads_11_917_p, data = ., na.action = na.omit)$statistic
  )

# View results
print(cbcl_t_tests)

# Format results for the table
table_data <- merged_diagnosis %>%
  summarise(
    `DSM-5 Anxiety` = paste0(
      round(mean(cbcl_scr_dsm5_anxdisord_t, na.rm = TRUE), 2), " (",
      round(sd(cbcl_scr_dsm5_anxdisord_t, na.rm = TRUE), 2), ")"
    ),
    `DSM-5 Depression` = paste0(
      round(mean(cbcl_scr_dsm5_depress_t, na.rm = TRUE), 2), " (",
      round(sd(cbcl_scr_dsm5_depress_t, na.rm = TRUE), 2), ")"
    )
  )

# Generate the Table Using gt
summary_data %>%
  gt() %>%
  tab_header(
    title = "Demographic Characteristics of the Sample"
  ) %>%
  cols_label(
    Full_Sample = "Full Sample",
    `OCD Absent` = "OCD Absent",
    `OCD Present` = "OCD Present",
    `Group Difference` = "Group Difference",
    p = "p",
    `Effect Size` = "Effect Size"
  ) %>%
  fmt_number(
    decimals = 2
  )
# Print the table
print(final_table)
```

#	OCS score
obsessive compulsive score derived from CBCL items
```{r}
ocs_score <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/ocd/p_data.csv")

ocs_score <- ocs_score %>%
  select(src_subject_id, ocs_score)
```

# Health variables
```{r}
health <- read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/physical-health/ph_p_mhx.csv") 

health <- health %>%
  select(src_subject_id, eventname, medhx_6i, medhx_6i_l, medhx_2h, medhx_6p, medhx_2h_l) %>%
  filter(eventname %in% c("baseline_year_1_arm_1", "2_year_follow_up_y_arm_1")) %>%
  mutate(
    head_injury = case_when(
      medhx_6i == 1 | medhx_6i_l == 1 ~ 1,
      TRUE ~ 0
    ),
    seizure = case_when(
      medhx_2h == 1 | medhx_6p == 1 | medhx_2h_l == 1 ~ 1,
      TRUE ~ 0
    )
  ) %>%
  select(src_subject_id, head_injury, seizure)
```

# Medication use
```{r}
medication <-  read.csv("/Users/maggieheimvik/Desktop/LCBC/Data/ABCD/core/physical-health/ph_p_meds.csv")

# Recode and simplify medication
```

